Truthfulness Verification System

### 2011-03-30 trunk/proposal/paper.tex
updated summery
Draft for Mar 30

{\em Wed Mar  9 09:30:15 CST 2011}

@@

@@

Moreover, we can also look for truthful statement by using those above extracted sentences. Using a better POS tagger and Named entity extractor (may those offered by LingPipe or Standford Named Entity Tagger) we cant extract the noun phrases and create the list of possible $AU$ that matches the data type of the $DU$. A comparison with the $AU$-s generated by the exisiting version of the T-verifier will give us allow us to do a fine grained analysis.

@@

{\em Feb 23 to Mar 9} 
\begin{itemize}
\item Formulated the procedure of truthfulness verification in a more structured way
\item As a result of the procedure formulation, the sections Step-by-step Algorithm for Truthfulness and Verification Examples of Truthfulness Verification are written in a complete revised way compared to the proposal
\item Apart from getting out feet wet with the MQL api, the daunting task is to develop generalized framework that is flexible enough to fit the wide variety of sentences into the rigid structure of Freebase. The learning cure is steep, given the richness and massive size of documentation. Though we are groping in dark currently, we are hopeful that the pieces of the puzzle with soon fit in.
\item Got some initial promising results by exploring beyond ontologies. 
\end{itemize}

\section{Conclusion}
It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.

>>>>

@@

{\em Wed Mar 9 2011}

@@

%Moreover, we can also look for truthful statement by using those above extracted sentences. Using a better POS tagger and Named entity extractor (may those offered by LingPipe or Standford Named Entity Tagger) we cant extract the noun phrases and create the list of possible $AU$ that matches the data type of the $DU$. A comparison with the $AU$-s generated by the exisiting version of the T-verifier will give us allow us to do a fine grained analysis.

Contrary to our prior report, Wikipedia does have a very rich API, which the wiki software, Mediawiki provides. Of the various features that this API provides there are two searching mechanisms.

\subsubsection{Opensearch Protocol}
The first one is Opensearch protocol, which gets pages whose name case-insensitively match a given string. When default limit(10) is reached, results are ordered by number of incoming links. For example if we search for Forrest Gump, we will get:

[

    * "Forrest Gump"
    *
      -
      [
          o "Forrest Gump"
          o "Forrest Gump (character)"
          o "Forrest Gump (novel)"
          o "Forrest Gump (soundtrack)"
          o "Forrest Gump – Original Motion Picture Score"
          o "Forrest Gump (disambiguation)"
      ]

]

This to a certain extent provides some semantic information attached alongwith the term.

\subsubsection{Fulltext search}
This provides a richer resultset, which syntactically resembles as the results returned for query on a web search engine. However, the interpretation of the returned results and the methods to analyse it has to significantly different. A real world example might make this distinction more clear. Searching the web is more like asking which books in the library has information on a particular topic, but full text search on wikipedia is more like pulling out a single volume of one's favourite encylopedia and searching the index pages to find if there is an article on this topic. So each article has a definite focus and it is closely defined by its title. Majority of the wikipedia pages adhere to some structure, which might reveal more than the text that is in there for the article. For example the interwiki links, the See also links etc. point towards related topics. Before exploring those areas, we first focus on the various aspects that are immediately available from the api:

\begin{enumerate}
\item    srinfo – What metadata to return. Type: one of totalhits, suggestion
\item    srlimit – How many total pages to return. Type: limit
\item    srnamespace – The namespace(s) to enumerate. Type: namespace
\item    sroffset – Use this value to continue paging (return by query). Type: integer
\item    srprop – What properties to return:
\begin{enumerate}
\item    size - Adds the size of the page in bytes
\item    wordcount - Adds the word count of the page
\item    timestamp - Adds the timestamp of when the page was last edited
\item    score - Adds the score (if any) from the search engine
\item    snippet - Adds a parsed snippet of the page
\item    titlesnippet - Adds a parsed snippet of the page title
\item    redirectsnippet - Adds a parsed snippet of the redirect
\item    redirecttitle - Adds a parsed snippet of the redirect title
\item    sectionsnippet - Adds a parsed snippet of the matching section
\item    sectiontitle - Adds a parsed snippet of the matching section title
\item    hasrelated - Indicates whether a related search is available
\end{enumerate}
\item    srredirects – Include redirect pages in the search. Type: bool
\item    srsearch (required) – Search for all page titles (or content) that has this value. Type: string
\item    srwhat – Search inside the text. Searching titles is disabled in Wikipedia.
\end{enumerate}

\subsection{Problem transformation - Recognizing Text Entailment}
While a thorough experimentation needs to be done in how to utlize all these parameters to do a better extraction of the underlying semantics, we realize limiting the knowledge base to Wikipedia, reduces the problem to an instance of Recognizing Text Entailment (RTE). Over the last few years amazing progress has been made in this track, and the state of the art algorithms are quite complicated to be implemented with in this short time frame. We start off with a simple idea for RTE, and list some experimentation that we plan to do ahead.

The main problem of RTE is that given a Hypothesis(H) and Text(T), the algorithm has to detect if T entails H.

T:The sale was made to pay Yukos' US\$ 27.5 billion tax bill, Yuganskneftegaz was originally sold for US\$ 9.4 billion to a little known company Baikalfinansgroup which was later bought by the Russian state-owned oil company Rosneft .

H:Baikalfinansgroup was sold to Rosneft.

The above example is taken from Recognizing Textual Entailment (RTE) 3 Challenge Corpora. The correct answer in this case, (i.e. yes T entails H) is easy for humans to comprehend, but obiviously the greatest challenging task to do programmatically.

The basis of the transformation is that each $DS$ now becomes a Hypothesis(H) and the wikipedia article the Text(T).

\subsection{With or without the doubt unit}

The original approach left out the doubt unit in order to not bias the returned results. However, in this case it might actually be beneficial. We can also go the other way, that is using the $DU$ to search article and then check in that article if there is sufficient intersection with the remainder of the doubtful statement. While this may work well for names and places, this would not work out for dates enforcing us to have separate rules attached to different data types. We have only tried without the DU right now.

\subsubsection{Without DU}
We remove the doubt unit (DU) and stop words from the topic unit and use the remaining key words to do a full text search. The results returned by Lucene, search engine that powers these searches are very accurate. But our task is much more complex, and it requires even finer analysis of the result.

Of the various properties that a returned result may be decorated with, the two most important ones are:
\begin{itemize}
\item titlesnippet - This is probably the most important category of all. If the doubt unit is located here then it would mean that there is a wikipedia page for the doubt unit, and the text contains keywords from the doubtful statement. This is the strongest indication that that the statement might be true.
\item snippet - This is the SRR in this case, that is formed by the concatenation of parts from three sentences that surround matched query keywords in the article content. The important thing to consider over here are the distaces between two such sentence parts and the number of matches each part includes. For a long page, there may be located at two very different areas in the page and be about very different contexts. In order to handle this we can take a weighted distance measure that takes the number of words between two matching sentence parts $0.33*number of matches in a sentence part/(length of the sentence part*length of document)$

\end{itemize}

			snippet_split_weights = {}
        		snippet_splits = split snippets by '<b>...</b>' 
                        for snippet_split in snippet_splits:
                                snippet_split_len = length(snippet_split)
                                snippet_split_match_count = count number of matches in snippet_split
                                snippet_split_weights[snippet_splits.index(snippet_split)] = snippet_split_match_count/snippet_split_len
                        snippet_weight = 0
                        for snippet_split_weight in snippet_split_weights.values():
                                snippet_weight +=snippet_split_weight
                        snippet_weight= 0.33*snippet_weight

                        #TITLE
                        title_match_count = count all matches in result['titlesnippet']
                        title_len = length(result['title'])
                        title_weight = title_match_count/title_len

                        #TODO need to determine appropriate wieghts
                        overall_result_weight = (title_weight+snippet_weight)/2

                        if regex_tokenize:
                                tokenizer = RegexpTokenizer('([A-Z]\.)+|\w+|\$[\d\.]+')

                                snippet_tokens = set(tokenizer.tokenize(snippet))
                                title_tokens = set(tokenizer.tokenize(result['title']))
                                doubt_unit_tokens = set(tokenizer.tokenize(doubt_unit))
                        else:
                                snippet_tokens = set(lower case the snippet and tokenize by space)
                                title_tokens = set(lower case the title and tokenize by space)
                                doubt_unit_tokens = set(lower case the doubt_unit and tokenize by space)

                        found = False

                        if snippet_tokens.intersection(doubt_unit_tokens):
                                print "S",';',title_weight,';',snippet_weight,';',overall_result_weight
                                found = True
                        if title_tokens.intersection(doubt_unit_tokens):
                                print "T",';',title_weight,';',snippet_weight,';',overall_result_weight
                                found = True

As for the rest of the properties like redirectitles, redirectsnippets, sectiontitle, sectionsnippets did not return values that seem to be of immediate help. However, they made find their use in generating alternative units. We have not yet experimented with them yet.

Currently the code that we have written only checks for presence of the doubt unit in each of the above mentioned property, for the fifty sentences that were there in the original paper.

Following are a few high priority bugs we are working on

\begin{itemize}
\item Better matching: We tried two ways to match the DU-s in the reutrened results, one a conservative approach that would try to tokenize so that abbreviations like "U.S.A" and monetary amounts like "\$23.00" are kept as tokens. In the more relaxed approach the tokens were all lowercased and split by spaces. This obviously is the most crucial part of the entire process and it appears we should have separate strategy based on the data type of the doubt unit.
\item Building the classifier: We got the data from TREC 9, which is in a question answer format and have finished coverting 100 of the 711 sentences all to statement form with correct answers. Once we have some more sentences labeled we will use it to train a classifier from the features extracted above.
\end{itemize}

@@

{\em Feb 23 to Mar 9} 
\begin{itemize}
\item Formulated the procedure of truthfulness verification in a more structured way
\item As a result of the procedure formulation, the sections Step-by-step Algorithm for Truthfulness and Verification Examples of Truthfulness Verification are written in a complete revised way compared to the proposal
\item Apart from getting out feet wet with the MQL api, the daunting task is to develop generalized framework that is flexible enough to fit the wide variety of sentences into the rigid structure of Freebase. The learning cure is steep, given the richness and massive size of documentation. Though we are groping in dark currently, we are hopeful that the pieces of the puzzle with soon fit in.
\item Looked into applicability of POS tagger in the context of our problem and found it useful for constructing freebase queries
\item Got some initial promising results by exploring different options such as Wikipedia, answer.com beyond Freebase ontologies. 
\item Worked on porting of existing T-verifier code to linux. The original T-verifier code is for windows only and not generic enough at this moment. We also contacted with the student author about the problems we are facing to port the code to linux. We hope to resolve it soon.
\end{itemize}

{\em Mar 9 to Mar 29} 
\begin{itemize}
\item 
\end{itemize}
\section{Conclusion}
It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.
